---
title: "formattable as an htmlwidget"
author: "Kent Russell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{formattable as an htmlwidget}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`formattable` was originally designed to offer additional formatting to the `markdown` generated by the deliberately sparse `knitr::kable`.  This design limited `formattable` to only the context of an `Rmarkdown` document.  Other contexts, such as the R console or RStudio IDE, would only see the much less attractive `character` markdown output.

```{r results='markup'}
library(formattable)

as.character(formattable(head(mtcars,3)))
```

Fortunately, a new function `as.htmlwidget` uses [`markdown`](http://cran.r-project.org/web/packages/markdown/index.html) to easily convert a `formattable` object to an `htmlwidget`.  Once converted to an `htmlwidget`, a user in these other contexts can leverage the infrastructure of [`htmlwidgets`](http://htmlwidgets.org) to benefit from the final `HTML` output of `formattable`.


## Examples

*note, please run these in an `?interactive` environment such as the console or RStudio IDE to see the difference*

Let's explicitly convert a `formattable` to an `htmlwidget` to help us understand what is happening.

```{r}
as.htmlwidget( formattable( head(mtcars,3) ) )
```

Although our `htmlwidget` version looks virtually the same when generated from a `Rmd` document such as this vignette, a user will have a noticeably different experience than before with this code.


